Weak Lensing Shear Inference:A Hierarchical Bayesian Approach

### 2013-03-12 docs/write-up.tex
add formulas

\title{Weak Lensing Shear Inference:A Hierarchy Bayesian Approach}

\date{02/19/2013}
\maketitle
\begin{abstract}
We present a Hierarchy Bayesian method for shear inference in Weak
Lensing survey. With this method we are able to infer synchronously
cosmic shear at different sky patches as well as the intrinsic shape
distribution. We show that this Hierarchy Bayesian method is more
accurate than the traditional average method and (prove to a large
extent )it is an unbiased estimation of shear.

Key words:gravational lensing-methods;data analysis-methods;bayesian
inference 
\end{abstract}

@@

\begin{equation}
L(g,\alpha|\epsilon_{\ell})=P(\epsilon_{\ell}|g,\alpha)=P(\epsilon_{0}(g)|\alpha)*|\frac{\partial(|\epsilon_{0}|,\phi_{0})}{\partial(|\epsilon_{\ell}|,\phi_{\ell})}|
\end{equation}

$\epsilon_{0,}\epsilon_{\ell,}$ means the unlensed and lensed elliticity.
$\alpha$ the parameter of the distribution of unlensed ellipticity.

Due to the homogeneous universe,we can believe that all the galaxies
in different sky patches have the same intrinsic shape distribution.Therefore,in
order to get higher accuracy,what we do is to infere reduced shears
of a number of sky patches simultaneously with their common shape
parameter $\alpha$. 
\begin{equation}
L(\vec{g},p,q|\epsilon_{\ell})=\prod_{i}\prod_{j}P(\epsilon_{0-j}(g_{i})|\alpha)|\frac{\partial(x_{0-j},y_{0-j})}{\partial(x_{\ell-j},y_{\ell-j})}|)
\end{equation}

i denotes ith patch and j denotes jth galaxies in a patch.

In this toy model,we choose two models for the distribution of unlensed
ellipticity :One is Beta distribution model.In this model we assume
the intrinsic ellipticity distribution is a beta distribution with
unknown parameter p and q.,The other one is step function model.Here
we devide 2D plane into a number of bins,and probability on each bin
is a free parameter.In later section we will show that we can get
similar results from these two models.

$|\frac{\partial(|\epsilon_{0}|,\phi_{0})}{\partial(|\epsilon|,\phi)}|$
is the Jacobean Determinant term. 
\begin{equation}
|\frac{\partial(|\epsilon_{0}|,\phi_{0})}{\partial(|\epsilon_{\ell}|,\phi_{\ell})}|=|\frac{\partial(|\epsilon_{0}|,\phi_{0})}{\partial(x_{0},y_{0})}*\frac{\partial(x_{0},y_{0})}{\partial(x_{\ell},y_{\ell})}*\frac{\partial(x_{\ell},y_{\ell})}{\partial(|\epsilon_{\ell}|,\phi_{\ell})}|=\frac{|\epsilon_{\ell}|}{|\epsilon_{0}|}*|\frac{\partial(x_{0},y_{0})}{\partial(x_{\ell},y_{\ell})}|
\end{equation}

$(x_{0},y_{0})$ and $(x_{\ell},y_{\ell})$ are the Cartesian coordinates
of $\epsilon_{0}$ and $\epsilon_{\ell}$ on the complex plane.This
Jacobean term is numerically computaed.

@@

When $|\epsilon_{0}|\longmapsto0$ ,$\epsilon\approx\epsilon_{0}+g$.Therefore,$|\frac{\partial(x_{0},y_{0)}}{\partial(x,y)}|\mapsto1$.Thus
when $|\epsilon_{0}|\mapsto0$ the Jacobi term $|\frac{\partial(|\epsilon_{0}|,\phi_{0})}{\partial(|\epsilon|,\phi)}|\mapsto\infty$.

@@

for HB method the typical value is \textasciitilde{}0.17 While for
traditional averaging method it's \textasciitilde{}0.18.

@@

@@

\begin{tabular}{|c|c|c|c|}
\hline 
Step Model & Expect Value  & $\sigma$ & S/N\tabularnewline
\hline 
\hline 
m1  & -6.09{*}$10^{-4}$ & 0.00239 & 0.25\tabularnewline
\hline 
c1  & 1.82$*10^{-4}$  & 1.44$*10^{-4}$ & 1.26\tabularnewline
\hline 
m2  & 0.00178  & 0.00281 & 0.63\tabularnewline
\hline 
c2  & 2.978$*10^{-4}$  & 1.71{*}$10^{-4}$ & 1.74\tabularnewline
\hline 
\end{tabular}

>>>>

\title{Weak Lensing Shear Inference:A Hierarchical Bayesian Approach}

\date{02/27/2013}
\maketitle
\begin{abstract}
We present an Hierarchical Bayesian approach to infer the shear map
caused by Weak Lensing effect,by solving a toy weak lensing problem,in
which we set a uniform shear on each sky patch and we have galaxies'
ellipticity data either with no measurement error or a simple 2-D
Gaussian noise. This hierarchical inference involves learning the
parameters of an exceedingly flexible model for the distribution of
unlensed galaxies' ellipticity, subject to a constraint that this
unlensed distribution is isotropical on ellipticity orientaion.It
is showed that using this flexible model P($\epsilon$|$\alpha$),we
are able to infer shears of different sky patches and their common
unlensed ellipticity distribution simultaneously.Though the average
of galaxies' ellipticity gives a simple estimation of cosmological
shear,we show that this hierarchical Bayesian method is more accurate
and precise.

Key words:gravational lensing-methods;data analysis-methods;hierarchical
bayesian inference 
\end{abstract}

@@

\begin{equation}
L(\vec{g},\alpha|\epsilon_{\ell})=p(\epsilon_{\ell}|\vec{g},\alpha)=P(\epsilon_{0}(\vec{g})|\alpha)*|\frac{\partial\epsilon_{0}}{\partial\epsilon_{\ell}}|
\end{equation}

Note that both $\epsilon_{0}$ and $\epsilon_{\ell}$are complex numbers.So
$|\frac{\partial\epsilon_{0}}{\partial\epsilon_{\ell}}|$ is the Jacobian
determinant term.Here the prior distribtuion we use for $\vec{g}$
is a 2Dflat distribution in the region $|g|<1$.Therefore $prior(\vec{g)}=\begin{cases}
1 & |\vec{g}|<1\\
0 & |\vec{g}|>1
\end{cases}$

The prior of $\alpha$ some how depends on the model we choose for
intrinsic ellipticity distribution .When we use the step function
model for P($\epsilon_{0}$),we find that we need to use a prior on
$\alpha$ .
\begin{equation}
Prior(\vec{\alpha}|\xi)=exp(-\xi\sum(\alpha_{i}-\alpha_{i-1})^{2})
\end{equation}

Given the likehood functions and priors we can also calculated the
marginalized likelihood function.
\begin{equation}
p(\epsilon_{\ell}|\vec{g})=\int p(\epsilon_{\ell}|\vec{g},\alpha)p(\alpha|\vec{g})d\alpha
\end{equation}

Since $p(\alpha)$ does not explicitly depends on $\vec{g}$ .$p(\alpha|\vec{g})$is
just $Prior(\vec{\alpha}|\xi)$.Therefore,we have marginalized likehood
\begin{equation}
p(\epsilon_{\ell}|\vec{g})=\int p(\epsilon_{\ell}|\vec{g},\alpha)Prior(\vec{\alpha}|\xi)d\alpha
\end{equation}

@@

When $|\epsilon_{0}|\longmapsto0$ ,$\epsilon\approx\epsilon_{0}+g$.Therefore,$|\frac{\partial(x_{0},y_{0)}}{\partial(x,y)}|\mapsto1$.Thus
when $|\epsilon_{0}|\mapsto0$ the Jacobean term $|\frac{\partial(|\epsilon_{0}|,\phi_{0})}{\partial(|\epsilon|,\phi)}|\mapsto\infty$.

@@

@@

\includegraphics[scale=0.5]{errorN8}

@@

\begin{tabular}{|c|c|c|c|}
\hline 
Step Model & Expect Value  & $\sigma$ & S/N\tabularnewline
\hline 
\hline 
m1  & -6.09{*}$10^{-4}$ & 0.00239 & 0.25\tabularnewline
\hline 
c1  & 1.82$*10^{-4}$  & 1.44$*10^{-4}$ & 1.26\tabularnewline
\hline 
m2  & -3.48{*}$10^{-4}$ & 0.00214 & 0.16\tabularnewline
\hline 
c2  & 2.01$*10^{-4}$  & 1.29{*}$10^{-4}$ & 1.56\tabularnewline
\hline 
\end{tabular}
